# Example

The following example does some basic plotting with R.

## Basic Code Expressions

Here is an example of inline code: 2 + 2 = `r 2 + 2`

Here is an example R codeblock with `echo=TRUE` (the default):

```{r}
2 + 2
```

Here is the same example R codeblock but with `echo=FALSE`:

```{r echo=FALSE}
2 + 2
```

Here is a sample of plotting data:

```{r}
head(cars)
plot(cars, xlab = "Speed (mph)", ylab = "Stopping distance (ft)", las = 1)
```

Finally, variables declared in one code block:

```{r}
x = 12
```

... can be referenced in another:

```{r}
x
```

This allows you to explain your analysis as you build it.


## Simple Analysis Example

For the followig analysis, we want to explore the `iris` package:

```{r}
head(iris)
```

Let's look at a plot of the data color-coded by species:

```{r}
plot(iris, col=iris$Species)
```

Finally, let's look at some scatterplots of the data (taken from a [question on Stack Overflow](http://stats.stackexchange.com/questions/30788/whats-a-good-way-to-use-r-to-make-a-scatterplot-that-separates-the-data-by-trea)):

Load the `ggplot` library:

```{r}
library(ggplot2)
```

Using this, we can see the data clustered by species:

```{r}
ggplot(iris, aes (x=Sepal.Length, y=Sepal.Width, colour=Species)) + stat_density2d()
```

And, as a reminder, if you didn't want your R code to be displayed when showing any part of this, you could simple include `echo=FALSE` in the `{r}` directives. This allows you to build up a complex series of instructions and just display the final output (the following is copied shamelessly from [Getting Fancy with 3-D Scatterplots](http://www.r-bloggers.com/getting-fancy-with-3-d-scatterplots/)).

Here's a 3-D scatterplot:

```{r echo=FALSE}
library(scatterplot3d)

with(mtcars, {
   scatterplot3d(disp,   # x axis
                 wt,     # y axis
                 mpg,    # z axis
                 main="3-D Scatterplot Example 1")
})
```

Here's the same scatterplot with vertical lines to help visualize the (X,Y) portions of the coordinates:

```{r echo=FALSE}
with(mtcars, {
   scatterplot3d(disp, wt, mpg,        # x y and z axis
                 color="blue", pch=19, # filled blue circles
                 type="h",             # lines to the horizontal plane
                 main="3-D Scatterplot Example 2",
                 xlab="Displacement (cu. in.)",
                 ylab="Weight (lb/1000)",
                 zlab="Miles/(US) Gallon")
})
```

Here's the same scatterplot with points labeled:

```{r echo=FALSE}
with(mtcars, {
   s3d <- scatterplot3d(disp, wt, mpg,        # x y and z axis
                 color="blue", pch=19,        # filled blue circles
                 type="h",                    # vertical lines to the x-y plane
                 main="3-D Scatterplot Example 3",
                 xlab="Displacement (cu. in.)",
                 ylab="Weight (lb/1000)",
                 zlab="Miles/(US) Gallon")
    s3d.coords <- s3d$xyz.convert(disp, wt, mpg) # convert 3D coords to 2D projection
    text(s3d.coords$x, s3d.coords$y,             # x and y coordinates
         labels=row.names(mtcars),               # text to plot
         cex=.5, pos=4)           # shrink text 50% and place to right of points)
})
```

And here's the same scatterplot with color-coding by class:

```{r echo=FALSE}
# create column indicating point color
mtcars$pcolor[mtcars$cyl==4] <- "red"
mtcars$pcolor[mtcars$cyl==6] <- "blue"
mtcars$pcolor[mtcars$cyl==8] <- "darkgreen"
with(mtcars, {
    s3d <- scatterplot3d(disp, wt, mpg,        # x y and z axis
                  color=pcolor, pch=19,        # circle color indicates no. of cylinders
                  type="h", lty.hplot=2,       # lines to the horizontal plane
                  scale.y=.75,                 # scale y axis (reduce by 25%)
                  main="3-D Scatterplot Example 4",
                  xlab="Displacement (cu. in.)",
                  ylab="Weight (lb/1000)",
                  zlab="Miles/(US) Gallon")
     s3d.coords <- s3d$xyz.convert(disp, wt, mpg)
     text(s3d.coords$x, s3d.coords$y,     # x and y coordinates
          labels=row.names(mtcars),       # text to plot
          pos=4, cex=.5)                  # shrink text 50% and place to right of points)
# add the legend
legend("topleft", inset=.05,      # location and inset
    bty="n", cex=.5,              # suppress legend box, shrink text 50%
    title="Number of Cylinders",
    c("4", "6", "8"), fill=c("red", "blue", "darkgreen"))
})
```

## Conclusion

The combination of R and Markdown makes for a very powerful report tool.
